A density ratio framework for evaluating the utility of synthetic data

### 2024-08-23 paper/paper.tex
Update paper

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[]{Latin Modern Roman}
  \setmathfont[]{Latin Modern Math}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\newcommand{\Xobs}{\mathbf{X}_{\text{obs}}}
\newcommand{\Xsyn}{\mathbf{X}_{\text{syn}}}
\newcommand{\dist}{\mathcal{P}}
\newcommand{\pobs}{p_{\text{obs}}}
\newcommand{\psyn}{p_{\text{syn}}}
\newcommand{\nobs}{n_{\text{obs}}}
\newcommand{\nsyn}{n_{\text{syn}}}
\newcommand{\bx}{\mathbf{x}}
\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A density ratio framework for evaluating the utility of synthetic data},
  pdfauthor={Thom Benjamin Volker; Peter-Paul de Wolf; Erik-Jan van Kesteren},
  pdfkeywords={synthetic data, utility, density ratio, privacy,
disclosure limitation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\usepackage{setspace}
\doublespacing
\newcommand{\runninghead}{A Preprint }
\renewcommand{\runninghead}{Density ratios for utility }
\title{A density ratio framework for evaluating the utility of synthetic
data}
\def\asep{\\\\\\ } % default: all authors on same column
\def\asep{\And }
\author{\textbf{Thom Benjamin
Volker}~\orcidlink{0000-0002-2408-7820}\\Methodology and Statistics
\textbar{} Methodology\\Utrecht University \textbar{} Statistics
Netherlands\\Utrecht,\ 3584CH\\\href{mailto:t.b.volker@uu.nl}{t.b.volker@uu.nl}\asep\textbf{Peter-Paul
de Wolf}\\Methodology\\Statistics Netherlands\\The
Hague,\ 2490HA\\\href{mailto:pp.dewolf@cbs.nl}{pp.dewolf@cbs.nl}\asep\textbf{Erik-Jan
van Kesteren}\\Methodology and Statistics\\Utrecht
University\\Utrecht,\ 3584CH\\\href{mailto:e.vankesteren1@uu.nl}{e.vankesteren1@uu.nl}}
\date{}
\begin{document}
\maketitle
\begin{abstract}
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO
\end{abstract}
{\bfseries \emph Keywords}
\def\sep{\textbullet\ }
synthetic data, utility, density ratio, privacy, disclosure
limitation \sep 
synthetic data, utility, density ratio, privacy, disclosure limitation

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, enhanced]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Openly accessible research data accelerates scientific progress
tremendously. Open data allows third-party researchers to answer
research questions with already collected data, freeing up resources
that would otherwise be devoted to data collection
(\protect\hyperlink{ref-ramachandran_open_2021}{Ramachandran, Bugbee,
and Murphy 2021}). Sharing data in combination with code allows others
to validate research findings and build upon the work
(\protect\hyperlink{ref-obels_analysis_2020}{Obels et al. 2020};
\protect\hyperlink{ref-crosas_automating_2015}{Crosas et al. 2015}).
Students can benefit from open data, as it fosters education with
realistic data (\protect\hyperlink{ref-atenas_open_2015}{Atenas,
Havemann, and Priego 2015}), as well as the general public, through
stimulating citizen science projects
(\protect\hyperlink{ref-newman_future_2012}{Newman et al. 2012}).
However, making data openly available is often (rightfully) hampered by
official legislation, like the General Data Protection Regulation (GDPR;
\protect\hyperlink{ref-gdpr}{European Parliament and Council of the
European Union 2016}), and general privacy concerns. In the worst case,
sharing data may cause harm to individuals or organizations, which may
withhold these entities from participating in future research. These
privacy constraints have been named among the biggest hurdles in the
advancement of computational social science
(\protect\hyperlink{ref-lazer_css_2009}{Lazer et al. 2009}), and among
top reasons for companies to not share their data with researchers
(\protect\hyperlink{ref-fpf_2017}{Future of Privacy Forum 2017}).

Multiple approaches exist to balance the benefits of open data with
potential privacy risks. Traditionally, data providers employed a suite
of different disclosure limitation techniques before sharing the data,
such as top-coding, record-swapping or adding noise (e.g.,
\protect\hyperlink{ref-hundepool_disclosure_2012}{Hundepool et al.
2012}; \protect\hyperlink{ref-willenborg_elements_2001}{Willenborg and
De Waal 2001}). More recently, synthetic data has gained traction as a
means to disseminate private data
(\protect\hyperlink{ref-SIPP_Beta_2006}{Abowd, Stinson, and Benedetto
2006}; \protect\hyperlink{ref-hawala_synthetic_2008}{Hawala 2008};
\protect\hyperlink{ref-drechsler2012}{Drechsler 2012};
\protect\hyperlink{ref-vandewiel2023}{van de Wiel et al. 2023};
\protect\hyperlink{ref-obermeyer2019}{Obermeyer et al. 2019};
\protect\hyperlink{ref-zettler2021}{Zettler et al. 2021}), although the
conceptual framework traces back to the previous century
(\protect\hyperlink{ref-little_statistical_1993}{Little 1993};
\protect\hyperlink{ref-rubin_statistical_1993}{Rubin 1993}). Simply put,
the idea of synthetic data is to replace some, or all, of the observed
values in a data set by synthetic values that are generated from some
model (e.g., \protect\hyperlink{ref-drechsler2011synthetic}{Drechsler
2011}). If only some values are replaced, disclosure risks can be
reduced because the sensitive or identifying values do not correspond to
their true values anymore. If all values are replaced, there is also no
one-to-one mapping between the original and the synthetic data, further
reducing the disclosure risk. However, an increase in privacy typically
comes at the cost of a decrease in utility. As more of the data is
altered, the quality of the released data becomes more sensitive to the
suitability of the generative model. Regardless of the approach to
disclosure limitation, if the technique used to generate or alter the
data does not align with the intricacies of the problem at hand, the
utility of the released data will be further reduced than necessary.

@@

In the synthetic data field, three classes of utility measures have been
distinguished (see \protect\hyperlink{ref-drechsler2023}{Drechsler and
Haensch 2023} for a thorough review): fit-for-purpose measures,
analysis-specific utility measures and global utility measures.
Fit-for-purpose measures are often the first step in assessing the
quality of the synthetic data. These typically involve comparing the
univariate distributions of the observed and synthetic data (for example
using visualization techniques or goodness-of-fit measures). Although
these measures provide an initial impression of the quality of the
synthesis models used, this picture is by definition limited, because
only one or two variables are assessed at the same time. Hence, complex
relationships between variables will always be out of scope. Such
relationships may be captured by analysis-specific utility measures,
which quantify whether analyses on synthetic data provide results that
are comparable to results from the same analysis performed on the
observed data. These measures can, for example, evaluate how similar the
coefficients of a regression model are (e.g., using the confidence
interval overlap; \protect\hyperlink{ref-karr_utility_2006}{Karr et al.
2006}), or whether prediction models trained on the synthetic and
observed data perform comparably in terms of evaluation metrics.
However, analysis-specific utility generally does not carry over: high
specific utility for one analysis does not at all imply high utility for
another analysis. Since data providers typically do not know which
analyses will be performed with the synthetic data, it is impossible to
provide analysis-specific utility measures for all potentially relevant
analyses (see also
\protect\hyperlink{ref-drechsler_utility_2022}{Drechsler 2022}).

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
    \setmainfont[]{Latin Modern Roman}
  \setmathfont[]{Latin Modern Math}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\newcommand{\Xobs}{\mathbf{X}_{\text{obs}}}
\newcommand{\Xsyn}{\mathbf{X}_{\text{syn}}}
\newcommand{\dist}{\mathcal{P}}
\newcommand{\pobs}{p_{\text{obs}}}
\newcommand{\psyn}{p_{\text{syn}}}
\newcommand{\nobs}{n_{\text{obs}}}
\newcommand{\nsyn}{n_{\text{syn}}}
\newcommand{\bx}{\mathbf{x}}
\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{float}
\floatplacement{table}{t}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A density ratio framework for evaluating the utility of synthetic data},
  pdfauthor={Thom Benjamin Volker; Peter-Paul de Wolf; Erik-Jan van Kesteren},
  pdfkeywords={synthetic data, utility, density ratio, privacy,
disclosure limitation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\usepackage{setspace}
\doublespacing
\newcommand{\runninghead}{A Preprint }
\renewcommand{\runninghead}{Density ratios for utility }
\title{A density ratio framework for evaluating the utility of synthetic
data}
\def\asep{\\\\\\ } % default: all authors on same column
\def\asep{\And }
\author{\textbf{Thom Benjamin
Volker}~\orcidlink{0000-0002-2408-7820}\\Methodology and Statistics
\textbar{} Methodology\\Utrecht University \textbar{} Statistics
Netherlands\\Utrecht,\ 3584CH\\\href{mailto:t.b.volker@uu.nl}{t.b.volker@uu.nl}\asep\textbf{Peter-Paul
de Wolf}\\Methodology\\Statistics Netherlands\\The
Hague,\ 2490HA\\\href{mailto:pp.dewolf@cbs.nl}{pp.dewolf@cbs.nl}\asep\textbf{Erik-Jan
van Kesteren}\\Methodology and Statistics\\Utrecht
University\\Utrecht,\ 3584CH\\\href{mailto:e.vankesteren1@uu.nl}{e.vankesteren1@uu.nl}}
\date{}
\begin{document}
\maketitle
\begin{abstract}
Synthetic data generation is a promising technique to facilitate the use
of sensitive data while mitigating the risk of privacy breaches.
However, for synthetic data to be useful in downstream analysis tasks,
it needs to be of sufficient quality. Various methods have been proposed
to measure the utility of synthetic data, but their results are often
incomplete or even misleading. In this paper, we propose using density
ratio estimation to improve quality evaluation for synthetic data, and
thereby the quality of synthesized datasets. We show how this framework
relates to and builds on existing measures, yielding global and local
utility measures that are informative and easy to interpret. We develop
an estimator which requires little to no manual tuning due to automatic
selection of a nonparametric density ratio model. Through simulations,
we find that density ratio estimation yields more accurate estimates of
global utility than established procedures. A real-world data
application demonstrates how the density ratio can guide refinements of
synthesis models and can be used to improve downstream analyses. We
conclude that density ratio estimation is a valuable tool in synthetic
data generation workflows and provide these methods in the accessible
open source R-package densityratio.
\end{abstract}
{\bfseries \emph Keywords}
\def\sep{\textbullet\ }
synthetic data, utility, density ratio, privacy, disclosure
limitation \sep 
synthetic data, utility, density ratio, privacy, disclosure limitation

\section{Introduction}\label{introduction}

\linespread{1}

\linespread{2}

Openly accessible research data accelerates scientific progress
tremendously. Open data allows third-party researchers to answer
research questions with already collected data, freeing up resources
that would otherwise be devoted to data collection
(\citeproc{ref-ramachandran_open_2021}{Ramachandran, Bugbee, and Murphy
2021}). Sharing data in combination with code allows others to validate
research findings and build upon the work
(\citeproc{ref-obels_analysis_2020}{Obels et al. 2020};
\citeproc{ref-crosas_automating_2015}{Crosas et al. 2015}). Students can
benefit from open data, as it fosters education with realistic data
(\citeproc{ref-atenas_open_2015}{Atenas, Havemann, and Priego 2015}), as
well as the general public, through stimulating citizen science projects
(\citeproc{ref-newman_future_2012}{Newman et al. 2012}). However, making
data openly available is often (rightfully) hampered by official
legislation, like the General Data Protection Regulation (GDPR;
\citeproc{ref-gdpr}{European Parliament and Council of the European
Union 2016}), and general privacy concerns. In the worst case, sharing
data may cause harm to individuals or organizations, which may withhold
these entities from participating in future research. These privacy
constraints have been named among the biggest hurdles in the advancement
of computational social science (\citeproc{ref-lazer_css_2009}{Lazer et
al. 2009}), and among top reasons for companies to not share their data
with researchers (\citeproc{ref-fpf_2017}{Future of Privacy Forum
2017}).

Multiple approaches exist to balance the benefits of open data with
potential privacy risks. Traditionally, data providers employed a suite
of different disclosure limitation techniques before sharing the data,
such as top-coding, record-swapping or adding noise (e.g.,
\citeproc{ref-hundepool_disclosure_2012}{Hundepool et al. 2012};
\citeproc{ref-willenborg_elements_2001}{Willenborg and De Waal 2001}).
More recently, synthetic data has gained traction as a means to
disseminate private data (\citeproc{ref-SIPP_Beta_2006}{Abowd, Stinson,
and Benedetto 2006}; \citeproc{ref-hawala_synthetic_2008}{Hawala 2008};
\citeproc{ref-drechsler2012}{Drechsler 2012};
\citeproc{ref-vandewiel2023}{van de Wiel et al. 2023};
\citeproc{ref-obermeyer2019}{Obermeyer et al. 2019};
\citeproc{ref-zettler2021}{Zettler et al. 2021}), although the
conceptual framework traces back to the previous century
(\citeproc{ref-little_statistical_1993}{Little 1993};
\citeproc{ref-rubin_statistical_1993}{Rubin 1993}). Simply put, the idea
of synthetic data is to replace some, or all, of the observed values in
a data set by synthetic values that are generated from some model (e.g.,
\citeproc{ref-drechsler2011synthetic}{Drechsler 2011}). If only some
values are replaced, disclosure risks can be reduced because the
sensitive or identifying values do not correspond to their true values
anymore. If all values are replaced, there is also no one-to-one mapping
between the original and the synthetic data, further reducing the
disclosure risk. However, an increase in privacy typically comes at the
cost of a decrease in utility. As more of the data is altered, the
quality of the released data becomes more sensitive to the suitability
of the generative model. Regardless of the approach to disclosure
limitation, if the technique used to generate or alter the data does not
align with the intricacies of the problem at hand, the utility of the
released data will be further reduced than necessary.

@@

In the synthetic data field, three classes of utility measures have been
distinguished (see \citeproc{ref-drechsler2023}{Drechsler and Haensch
2023} for a thorough review): fit-for-purpose measures,
analysis-specific utility measures and global utility measures.
Fit-for-purpose measures are often the first step in assessing the
quality of the synthetic data. These typically involve comparing the
univariate distributions of the observed and synthetic data (for example
using visualization techniques or goodness-of-fit measures). Although
these measures provide an initial impression of the quality of the
synthesis models used, this picture is by definition limited, because
only one or two variables are assessed at the same time. Hence, complex
relationships between variables will always be out of scope. Such
relationships may be captured by analysis-specific utility measures,
which quantify whether analyses on synthetic data provide results that
are comparable to results from the same analysis performed on the
observed data. These measures can, for example, evaluate how similar the
coefficients of a regression model are (e.g., using the confidence
interval overlap; \citeproc{ref-karr_utility_2006}{Karr et al. 2006}),
or whether prediction models trained on the synthetic and observed data
perform comparably in terms of evaluation metrics. However,
analysis-specific utility generally does not carry over: high specific
utility for one analysis does not at all imply high utility for another
analysis. Since data providers typically do not know which analyses will
be performed with the synthetic data, it is impossible to provide
analysis-specific utility measures for all potentially relevant analyses
(see also \citeproc{ref-drechsler_utility_2022}{Drechsler 2022}).

Global utility measures may overcome the shortcomings of the previous
approaches, as they evaluate the discrepancy between the entire
multivariate distribution of the observed and synthetic data. As such,
global utility measures yields the most promising class of utility
measures, because if the observed and synthetic data have similar
(multivariate) distributions, all potential analyses should yield
similar results. Global utility can be evaluated using some divergence
measure, such as the Kullback-Leibler divergence
(\citeproc{ref-karr_utility_2006}{Karr et al. 2006}), or by evaluating
whether the observed and synthetic data are distinguishable using a
classification model (a technique called \(pMSE\);
\citeproc{ref-Woo_global_2009}{Woo et al. 2009};
\citeproc{ref-snoke_utility_2018}{Snoke et al. 2018}). However, a common
critique of global utility measures is that they tend to be too general
(\citeproc{ref-drechsler_utility_2022}{Drechsler 2022}). That is,
analyses on a synthetic data set that is overall quite similar to the
observed data (i.e., has high global utility), may still yield results
that are far from the results obtained from the real data. Also,
commonly used methods for estimating the \(pMSE\), as logistic
regression and classification and regression trees, tend to become less
reliable as the dimensionality of the data increases, and are vulnerable
to model misspecification
(\citeproc{ref-drechsler_utility_2022}{Drechsler 2022}). Lastly, the
output of global utility measures can be hard to interpret, and say
little about the regions in which the synthetic data do not resemble the
true data accurately enough.
